---
title: "PA1_template"
author: "Devara Izaz Fathan"
date: "8/17/2020"
output: html_document
---
#Add The Packages
Add the packages needed in this projcet
```{r}
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
```

#Read The Data
The file can be download from link in the this link [Activity Data](https://raw.githubusercontent.com/devarafathan/RepData_PeerAssessment1/master/activity.zip)

```{r, echo=TRUE}
getwd()
url <- 'https://raw.githubusercontent.com/devarafathan/RepData_PeerAssessment1/master/activity.zip'
dest <- download.file(url, 'E:/BELAJAR CODING/R/COURSERA/RepData_PeerAssessment1/activity.zip')
unzip('activity.zip')
activity <- read.csv('activity.csv')

```

#Process The Data
The first data that required on this project is data without imputing. Therefore, we have to igonore NA values in the acitvity dataset. Here are the code  
for cleaning up and processing the dataset.

```{r, echo=TRUE}
activity_clean <- activity %>% filter(steps != 'NA')
head(activity_clean)

```

Then we can grouped the data based on the date and then summarize it to take the sum value for each day.
```{r, echo=TRUE}
activity_clean$date <- as.factor(activity_clean$date)
activity_histogram <- activity_clean %>% group_by(date) %>% summarize(total= sum(steps))
head(activity_histogram, 10)
```

#Histogram of The Total Number of Steps Taken Each Day
We will use ggplot function to plot the histogram of the total number of steps taken each day. 

```{r, echo=TRUE}

activity_histogram %>% ggplot(aes(total)) +
      geom_histogram(binwidth = 5000, center= 2500, fill='blue', col='black') + 
      scale_x_continuous(breaks = seq(0, 25000, 5000)) +
      scale_y_continuous(breaks= seq(0,30, 5)) +
      ggtitle('Histogram of The Total Number of Steps Each Day') +
      xlab('Total') + 
      ylab('Frequency')

```

#Mean And Median Number of Steps Taken Each Day
Calculate the Mean and Median Number using the group_by() and summarize() function.

```{r, echo=TRUE}
activity_mean_median <- activity_clean %>% group_by(date) %>% summarize(mean=mean(steps), median=median(steps))
activity_mean_median

```

#Time Series Plot of The Average Number of Steps Taken
First, we have to group the activity_clean dataset based on the interval and summarize the mean for each interval. Save the last table to a variable named activity_series.

```{r}
activity_series <- activity_clean %>% group_by(interval) %>% summarize(average= mean(steps))
head(activity_series, 10)

```

Then just plot the time series plot using activity_series dataset.

```{r, echo=TRUE}
activity_series %>% ggplot(aes(interval, average)) + 
      geom_line() + 
      scale_x_continuous(breaks = seq(0, 2500, 200)) +
      ggtitle('Time Series Plot of The Average Number of Steps Taken') +
      xlab('Interval') + 
      ylab('Average Numbers of steps')


```

#The 5-minute interval that, on average, contains the maximum number of steps
Find the maximum interval that have the maximum number of steps.

```{r, echo =TRUE}
index_max <- which.max(activity_series$average)
activity_series[index_max,]

```
**The maximum average step is at interval 835 with 206.1698 value**

#Code to describe and show a strategy for imputing missing data
Note that there are a number of days/intervals where there are missing values (coded as \color{red}{\verb|NA|}NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

In this project i choose the strategy to fill the NA steps value using the average interval value. Here is the code:

```{r, echo=TRUE}
activity_imputed <- activity
for (i in 1:nrow(activity_imputed)) {
      if(is.na(activity_imputed$steps[i])){
            inter <- activity_imputed$interval[i]
            average <- activity_series$average[which(activity_series$interval==inter)]
            activity_imputed$steps[i] <- average
      }
}
head(activity_imputed, 10)

```

#Histogram of The Total Number of Steps Taken Each Day After Missing Values are Imputed
Make a histogram after filling the NA value, so we use activity_imputed dataset. We have to group the date and the summarize the total steps of each day
```{r, echo=TRUE}
#group and summarize
activity_histogram2 <- activity_imputed %>% group_by(date) %>% summarize(total=sum(steps))

#plot the histogram
activity_histogram2 %>% ggplot(aes(total)) +
      geom_histogram(binwidth = 5000, center= 2500, fill='blue', col='black') + 
      scale_x_continuous(breaks = seq(0, 25000, 5000)) +
      scale_y_continuous(breaks= seq(0,40, 5)) +
      ggtitle('Histogram of The Total Number of Steps Each Day(Imputed)') +
      xlab('Total') + 
      ylab('Frequency')

      
```
After imputing the data, we can compare the data before and after computing using median and mean value
```{r, echo=TRUE}
#Mean before imputing
mean(activity_histogram$total)
```

```{r, echo=TRUE}
#mean after imputing
mean(activity_histogram2$total)
```

```{r, echo=TRUE}
#median before imputing
median(activity_histogram$total)

```

```{r, echo=TRUE}
#median after imputing
median(activity_histogram2$total)
```
#Panel Plot Comparing The Average Number of Steps Taken per 5-minute Interval Across Weekdays and Weekends
For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

##Create New Variable Vector
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
#Convert column date into class of date variable.
activity_imputed$date <- weekdays(as.Date(activity_imputed$date))
weekend_name <- c('Saturday', 'Sunday')
weekday_name <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
activity_imputed <- activity_imputed %>% mutate(typeday= ifelse(date %in% weekday_name, 'Weekday','Weekend'))
head(activity_imputed)
```

##Create The Plot

```{r, echo=TRUE}
activity_week <- activity_imputed %>% group_by(typeday, interval) %>% summarize(mean=mean(steps)) 
activity_week %>% ggplot(aes(interval, mean)) + 
      geom_line() + 
      scale_x_continuous(breaks = seq(0, 2500, 500)) +
      ggtitle('Time Series Plot') +
      xlab('Interval') + 
      ylab('Numbers of steps') +
      facet_wrap(.~typeday, ncol = 1)

```





